```{r setup,include=FALSE}
project<-"analysisVariance"
mygit<-"~/Arbeitsfläche/linModel/"
setwd(paste(mygit,project,sep="/"))
library("ProjectTemplate")
load.project()
# set global chunk options
opts_chunk$set(cache=FALSE,dev = c('png',"pdf"),echo=FALSE, message=FALSE, fig.width=11,fig.height=8,warning=FALSE,results='hide')

```

Varianzanalyse im linearen Modell
========================================================
## data given page 666 in Statisik, Sachs
[lm intro](http://data.princeton.edu/R/linearModels.html "Princeton R intro")
```{r dataS666high,results='markup'}
A<-c(13.2,14.1,7.8,11.7)
B<-c(15.9,16.2,19.3,18.0,17.3)
C<-c(6.8,9.2,12.4)
exp1<-data.frame(diameter=c(A,B,C),drug=c(rep("A",length(A)), rep("B",length(B)),rep("C",length(C))),Conc="high")
sapply(exp1,class)
head(exp1)
```

# linear model
```{r lm,results='markup'}
fm1<-lm(diameter~drug,exp1)
anova(fm1)
summary(fm1)
summary(glht(fm1,linfct=c("drugB=0","drugC=0")))
confint(fm1)
K <- diag(3)
rownames(K) <- names(coef(fm1))

summary(glht(fm1,linfct=K))
contr.treatment(3,base=1,contrasts=TRUE)

model.matrix(fm1)
```
# multiple pairwise comparisons using glht

```{r glht,results='markup'}
amod<-aov(diameter~drug,exp1)
summary(amod)
amodSum<-summary(amod)
amodGlht<-glht(amod,linfct=mcp(drug="Tukey"))


amodSG<-summary(glht(amod,linfct=mcp(drug="Tukey")))
plot(amodGlht)
```



```{r dataS666low,results='markup'}
A<-c(10.4,12.6,6.3)
B<-c(11.5,13.7,10.9,15.1)
C<-c(12.3,14.5,16.7,10.3)
exp2<-data.frame(diameter=c(A,B,C),drug=c(rep("A",length(A)), rep("B",length(B)),rep("C",length(C))),Conc="low")

sapply(exp2,class)
head(exp2)

```

```{r fm1abc}
exp12<-rbind(exp1,exp2)
sapply(exp12,levels)
head(exp12)
with(exp12,interaction.plot(drug,Conc,diameter,las=1))
fm12<-lm(diameter~drug+Conc,exp12)
fm12Glht<-glht(fm12,linfct=mcp(drug="Tukey"))
plot(fm12Glht)
fm12a<-update(fm12,.~.+drug:Conc)
fm12b<-update(fm12a,.~.-Conc)
fm12c<-update(fm12a,.~.-Gene)
fm12d<-lm(diameter~Conc+Conc:drug,exp12)
fm12e<-lm(diameter~Conc*drug,exp12)
anova(fm12b,fm12a,fm12)
fm<-fm12a
```
# compare models * and :
```{r fma,results='markup'}
fmGlht<-glht(fm,linfct=mcp(drug="Tukey"))
plot(fmGlht)
summary(fmGlht)
anova(fm)
summary(fm)
```
```{r fm12b}
fm<-fm12b
```

```{r fmb,ref.label="fma",results='markup'}

```

```{r fm12c}
fm<-fm12c
```

```{r fmc,ref.label="fma",results='markup'}

```


```{r fmGlht}
#cbind(exp12,IPRED=fitted(fm12a))
unique(fitted(fm12a))
#summary(glht(fm12a,linfct=c("drugB=0","drugC=0","drug:Conc"=0)))
K <- diag(6)
rownames(K) <- names(coef(fm12a))

summary(glht(fm12b,linfct=K[-1,]))
summary(glht(fm12b,linfct=K[-(1:3),]))
fitted(fm12a)
coef(fm12a)
summary(glht(fm12a))
```


```{r}
fm<-fm12a
unique(fitted(fm))
co<-coef(fm)
AH<-co[1]
BHmAH<-co[2]
CHmAH<-co[3]
ALmAH<-co[4]
BLmBHmALmAH<-co[5]
CLmCHmALmAH<-co[6]

c(AH=AH,BH=AH+BHmAH,CH=AH+CHmAH,AL=AH+ALmAH,BL=AH+BHmAH+ALmAH+BLmBHmALmAH,CL=AH+CHmAH+ALmAH+CLmCHmALmAH)

fm<-fm12b
unique(fitted(fm))
co<-coef(fm)
AH<-co[1]
BHmAH<-co[2]
CHmAH<-co[3]
ALmAH<-co[4]
BLmBH<-co[5]
CLmCH<-co[6]

c(AH=AH,BH=AH+BHmAH,CH=AH+CHmAH,AL=AH+ALmAH,BL=AH+BHmAH+BLmBH,CL=AH+CHmAH+CLmCH)
```


- AH=AH
- BH=AH+BHmAH
- CH=AH+CHmAH
- AL=AH+ALmAH
- BL=AH+BHmAH+ALmAH+BLmBHmALmAH
- CL=AH+CHmAH+ALmAH+CLmCHmALmAH)

BL-BH-(AL-AH)= BL-BH-AL+AH=(BL-AL)-(BH-AH)

A=log(Ref)
B=log(Gene)
(BL-AL)-(BH-AH) = log(Gene/Ref)L - log(Gene/Ref)H =log((Gene/Ref)L /(Gene/Ref)H)

fold increase of Gene/Ref by changing from H to L


- AH=AH
- BH=AH+BHmAH
- CH=AH+CHmAH
- AL=AH+ALmAH
- BL=BH+BLmBH= AH+BHmAH +BLmBH = AH+BH-AH + BL-BH
- CL=AH+CHmAH+ALmAH+CLmCHmALmAH)

c0[4]=AL-AH = log(RefL/RefH)
c0[5]=BL-BH = BL-BH + AH -AH = (BL-AH)-(BH-AH)= log((GeneL/RefH) /(Gene/Ref)H)
c0[6]=CL-CH

fold increase of Gene  by changing from H to L

Gene 



# Conclusion

- the difference between diameter A high Conc vs  diameter A low Conc is `r ALmAH`
- the difference between diameter B high Conc vs  diameter B low Conc is `r ALmAH+BLmBHmALmAH`  
 - drug B adds a difference of `r BLmBHmALmAH` to the A difference `r ALmAH`
- the difference between diameter C high Conc vs  diameter C low Conc is `r ALmAH+CLmCHmALmAH`
 - drug C adds a difference of `r CLmCHmALmAH` to the A difference `r ALmAH`


- Because in xALD all gene data is normalized to the reference gene HPRT or GAPDH, we want to know the fold increase of the Gene/HPRT ratio between the conc of 2,5,7 mol and 0 mol.
